---
title: "Community Health Outreach"
subtitle: ""
author: 
  name: "John Ryan Kivela, MA"
  email: "Ryan.Kivela@narbha.org"
  affiliation: "The Alliance ACO"
date: today
date-format: long
format:
  html:
    theme: pulse
    embed-resources: true
    toc: true
    toc-depth: 6
    code-fold: true
    footnotes-hover: true
---

This document creates the Community Health Outreach Roster. The CHO is a collection of enrollment data, demographics, VBP results, and other indicators of clinical conditions and social risk factors. This roster will be used for ACO reporting, provider clinical outreach, and predictive modeling.

# Method

## r Setup

This sets up the R Studio with needed packages and other setting for the HTML rendering and data viewing.

```{r}
#| label: Setup
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

## Load Libraries
library(tidyverse)
library(readxl)
library(odbc)
library(stringr)

# Set up ODBC Connection
# QRPTP01 <- DBI::dbConnect(odbc::odbc(),"AllianceData")

# adjust to allow viewing of more columns
options(width = 999)
```

## Pre-load data for testing

These are datasets that are created in the Framework. They are written to CSV in the Framework, and can be imported directly here if they are not already present in the environment. 

```{r}
#| label: Pre-Load Data
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

AllianceEnrolledMembers <- read.csv("./data/input/AllianceEnrolledMembers.csv")

Claims_AllHCA <- read.csv("./data/input/Claims_AllHCA.csv")

# ClaimsAMM_AllHCA <- read.csv("./data/input/ClaimsAMM_AllHCA.csv")

# ClaimsPBM_AllHCA <- read.csv("./data/input/Claims_AllHCA.csv")

Data_VBPQR_AllRecords <- read.csv("./data/input/Data_VBPQR_AllRecords.csv")

InpatientCensus <- read.csv("./data/input/AllianceDailyInpatientCensus_All2023.csv")

HHPrescriberNPI <- read.csv("./documents/HHPrescriberNPI.csv")

```

## Common Objects

Common objects are used throughout the program for queries and analysis.

```{r}
#| label: Create Common Objects
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Create table for inline code
# InLineCode <- data.frame(
#   Column1 = "text")

# Vector of Alliance Provider Name
Provider_ShortName <- (c("CBI", "CPIH", "EHS", "LCBHC", "MMHC", "SHG", "SBH", "TGC", "WYGC"))

# Vector of NAZ Counties
NAZ_Counties <- (c("Apache", "Coconino", "Mohave", "Navajo", "Yavapai"))

# Create provider shortnames based on TIN
# Create a data frame called ProviderShortnames
ProviderTIN <- 
  data.frame(
  BHHProvider_TIN = c("86-0206928", "86-0206928", "86-0207499", "86-0214457", "86-0215065", "86-0223720", "86-0250938", "86-0290033", "86-0473221", "94-2880847"),
  BHHShortName = c("PH", "WYGC", "SHG", "MMHC", "CPIH", "TGC", "LCBHC", "SBH", "EHS", "CBI")
  )

# Create a list with the ultimate column names for the data
VBPTrendReport_colnames <- 
  c("ClaimsAdjudicatedThrough",
    "AllianceProviderTIN",
    "ProviderShortname",
    "HealthHomeName",
    "MemberID",
    "LastName",
    "FirstName",
    "MeasureID",
    "Measure",
    "GapStatus")

# Contract Measures List
ContractMeasures <- c("AMM", "FUH", "FUM", "HDO", "FUA")

# HEDIS Value Sets
MyHEDIS2023_AllValueSetsToCodes <- read_xlsx("./documents/M. HEDIS MY 2023 Volume 2 Value Set Directory 2023-03-31 - Copy.xlsx", sheet = "Value Sets to Codes")

MyHEDIS2023_MedicationDirectory <- read_xlsx("./documents/HEDIS MY 2023 Medication List Directory 2023-03-31.xlsx", sheet = "Medication Lists to Codes")

# Z Codes
Z_Codes <- read_xlsx("./documents/Z-Codes.xlsx", sheet = "Sheet1")

```

## Procedure

### Enrollment Roster

```{r}
#| label: Enrollment roster
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

AllianceEnrolledMembers <-
  AllianceEnrolledMembers |> 
    mutate(
    EnrollmentDate = as.Date(EnrollmentDate, format = "%Y-%m-%d"),
    DisenrollmentDate = as.Date(DisenrollmentDate, format = "%Y-%m-%d"),
    DaysPerEnrollment = as.numeric(difftime(DisenrollmentDate, EnrollmentDate, units = "days"))
  )

```

### VBP Quality Roster

Create an indicator for if member is on VBPQR and add it to the Enrollment Roster

```{r}
#| label: Create VBPQR Indicator
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Select variables for analysis from VBPQR
VBPQR_filtered <-
  Data_VBPQR_AllRecords |> 
  select(
    ClaimsAdjudicatedThrough,
    AllianceProviderTIN,
    ProviderShortname,
    MemberID,
    MeasureID,
    GapStatus
  ) |> 
  rename("AHCCCSID" = MemberID) |> 
  filter(MeasureID != "HDO")

# Format as date
VBPQR_filtered$ClaimsAdjudicatedThrough <- as.Date(VBPQR_filtered$ClaimsAdjudicatedThrough)

# Step 1: Filter for only the most recent ClaimsAdjudicatedThrough date
VBPQR_filtered <- VBPQR_filtered |>
  group_by(AHCCCSID) |>
  filter(ClaimsAdjudicatedThrough == max(ClaimsAdjudicatedThrough))

# Step 2: Create new columns "AMM", "FUH", and "FUM". This is where the VBPQR indiciator will go.
VBPQR_filtered <- VBPQR_filtered |>
  mutate(AMM = NA, FUH = NA, FUM = NA)

# Step 3: Update new columns based on MeasureID. 
VBPQR_filtered$AMM[VBPQR_filtered$MeasureID == "AMM"] <- VBPQR_filtered$GapStatus[VBPQR_filtered$MeasureID == "AMM"]

VBPQR_filtered$FUH[VBPQR_filtered$MeasureID == "FUH"] <- VBPQR_filtered$GapStatus[VBPQR_filtered$MeasureID == "FUH"]

VBPQR_filtered$FUM[VBPQR_filtered$MeasureID == "FUM"] <- VBPQR_filtered$GapStatus[VBPQR_filtered$MeasureID == "FUM"]

# Collapse data to one row per AHCCCSID with an indicator for each measure.
collapsed_data <- VBPQR_filtered |>
  group_by(AHCCCSID) |>
  summarize(
    AMM = max(AMM, na.rm = TRUE),
    FUH = max(FUH, na.rm = TRUE),
    FUM = max(FUM, na.rm = TRUE)
  ) |> 
  mutate(AMM_Ind = if_else(!is.na(AMM), 1, 0),
         FUH_Ind = if_else(!is.na(FUH), 1, 0),
         FUM_Ind = if_else(!is.na(FUM), 1, 0)
         )

# Add the total numbers of VBP measures
collapsed_data <-
  collapsed_data |> 
  mutate(TotalVBP = AMM_Ind + FUH_Ind + FUM_Ind)

# Rename AHCCCSID for enrollment roster
AllianceEnrolledMembers <-
  AllianceEnrolledMembers |> 
    rename(AHCCCSID = "AhcccsId")

# Attach the VBPQR indicators to the enrollment roster by AHCCCSID 
CHO_Roster1 <-
  merge(x = AllianceEnrolledMembers,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

# Check it and see, you're hot blooded :)
Checkit <- CHO_Roster1 |> select(1, 36:43)
```

### Claims

Filter the AllHCA claims table for only those people on the Alliance Enrolled Members roster. Then attach the CHO Roster data to each claim. This data frame will be a foundation for creating many indicators.

```{r}
#| label: Select Claims
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Select variables to analyze from the claims data
Claims_AllAlliance <-
  Claims_AllHCA |> 
  filter(AHCCCSID %in% AllianceEnrolledMembers$AHCCCSID) |> # Filter for Alliance Members Only
  select(
    Icnno,
    PrimaryID,
    AHCCCSID,
    Age,
    #Age_Group,
    #Population,
    RA,
    Svccode,
    #Units,
    BegDate,
    PrimaryDiagnosis,
    ProviderName,
    Fedtaxid,
    RenderingProviderNpi,
    #PayContract,
    Placesvc,
    ProviderType,
    Dx1, Dx2, DX3, Dx4, Dx5, DX6,
    Dx7, Dx8, DX9, Dx10, Dx11, DX12,
    Calcnetpd,
    #MemberACCGSA,
    #MemberServiceArea,
    #MemberZipcode,
    #RenderingProviderServiceArea,
    MedicationName,
    MedicationGPI,
    MedicationNDC,
    #EncounterStatus,
    #EncounterStatusDate
  ) |> 
  rename(Dx3 = "DX3",
         Dx6 = "DX6",
         Dx9 = "DX9",
         Dx12 = "DX12")

# Remove any rogues spaces in any field in the data
Claims_AllAlliance <-
  Claims_AllAlliance |> 
  mutate_all(str_squish)

# merge the claims data with the CHO tool
CHO_Roster2 <-
  merge(x = Claims_AllAlliance,
        y = CHO_Roster1,
        by = "AHCCCSID",
        all = TRUE)

```

### Health Home Service Provision

This code provides an indicator for if the person has received services at either: 

  1.    Yes, at HH at least once in the last 6 months; 
  2.    No, never at their HH in the last 6 months: and 
  3.    Yes, the person has received services at both HH and non-HH.

Calculations for the first option include cases where the service provider was associated with an Alliance Provider, but used a different Tax Identification Number.

```{r}
#| label: Same prvider indicator
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Clean up Provider TIN data
# Remove any rogues spaces in any field in the data
HHPrescriberNPI <-
  HHPrescriberNPI |> 
  mutate_all(str_squish)

# Indicaters for services by the HH (Including HH prescribers using their own NPI numbers) or not by the HH
ProviderTINMatch <-
  CHO_Roster2 |> 
  mutate(SvcByAssignedBHHTIN = if_else(Fedtaxid == BHHProvider_TIN | RenderingProviderNpi == HHPrescriberNPI$NPI, 1, 0)) |> 
  mutate(SvcByOtherTIN = if_else(Fedtaxid != BHHProvider_TIN & RenderingProviderNpi != HHPrescriberNPI$NPI, 1, 0)) |> 
  select(1, 72, 73)

# Collapse data to a single row per person

# Group by AHCCCSID and summarize using max to check if 1 occurs anywhere
collapsed_data <- ProviderTINMatch |>
  group_by(AHCCCSID) |>
  summarize(
    SvcByAssignedBHHTIN = max(SvcByAssignedBHHTIN),
    SvcByOtherTIN = max(SvcByOtherTIN)
  )

collapsed_data <-
  collapsed_data |> 
  mutate(SvcByBoth = if_else(SvcByAssignedBHHTIN + SvcByOtherTIN == 2, 1, 0))

collapsed_data <-
  collapsed_data |> 
  mutate(NoSvcs = if_else(SvcByAssignedBHHTIN == 0 & SvcByOtherTIN == 0, 1, 0))

# Add new indicators to CHO_Roster1
CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

# Check it
Checkit <- CHO_Roster1 |> select(1, 36:47)

```

### Social Risk Factors 

These indicate if a person has a "Z-Code" diagnosis for each of several categories of social risk factors. In addition, this section provides a summary of the quantity of social risk factors experienced by a given individual.

#### All Z-Codes

```{r}
#| label: Z Code indicators
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Assuming CHO_Roster2 is your data frame
# Create new columns Z55, Z56, Z57, Z59, Z60, Z62, Z63, Z64, Z65, Z75
CHO_Roster2$Zcd_Z55 <- 0
CHO_Roster2$Zcd_Z56 <- 0
CHO_Roster2$Zcd_Z57 <- 0
CHO_Roster2$Zcd_Z59 <- 0
CHO_Roster2$Zcd_Z60 <- 0
CHO_Roster2$Zcd_Z62 <- 0
CHO_Roster2$Zcd_Z63 <- 0
CHO_Roster2$Zcd_Z64 <- 0
CHO_Roster2$Zcd_Z65 <- 0
CHO_Roster2$Zcd_Z75 <- 0

# Look through each column starting with Dx
for (col in names(CHO_Roster2)) {
  if (startsWith(col, "Dx")) {
    # Check for any value that starts with "Z55" and update Z55 column
    CHO_Roster2$Zcd_Z55 <- ifelse(grepl("^Z55", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z55)
    # Check for any value that starts with "Z56" and update Z56 column
    CHO_Roster2$Zcd_Z56 <- ifelse(grepl("^Z56", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z56)
    # Check for any value that starts with "Z57" and update Z57 column
    CHO_Roster2$Zcd_Z57 <- ifelse(grepl("^Z57", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z57)
    # Check for any value that starts with "Z59" and update Z59 column
    CHO_Roster2$Zcd_Z59 <- ifelse(grepl("^Z59", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z59)
    # Check for any value that starts with "Z60" and update Z60 column
    CHO_Roster2$Zcd_Z60 <- ifelse(grepl("^Z60", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z60)
    # Check for any value that starts with "Z62" and update Z62 column
    CHO_Roster2$Zcd_Z62 <- ifelse(grepl("^Z62", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z62)
    # Check for any value that starts with "Z63" and update Z63 column
    CHO_Roster2$Zcd_Z63 <- ifelse(grepl("^Z63", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z63)
    # Check for any value that starts with "Z64" and update Z64 column
    CHO_Roster2$Zcd_Z64 <- ifelse(grepl("^Z64", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z64)
    # Check for any value that starts with "Z65" and update Z65 column
    CHO_Roster2$Zcd_Z65 <- ifelse(grepl("^Z65", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z65)
    # Check for any value that starts with "Z75" and update Z75 column
    CHO_Roster2$Zcd_Z75 <- ifelse(grepl("^Z75", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z75)
  }
}

# Look through each column starting with PrimaryDiagnosis
for (col in names(CHO_Roster2)) {
  if (startsWith(col, "PrimaryDiagnosis")) {
    # Check for any value that starts with "Z55" and update Z55 column
    CHO_Roster2$Zcd_Z55 <- ifelse(grepl("^Z55", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z55)
    # Check for any value that starts with "Z56" and update Z56 column
    CHO_Roster2$Zcd_Z56 <- ifelse(grepl("^Z56", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z56)
    # Check for any value that starts with "Z57" and update Z57 column
    CHO_Roster2$Zcd_Z57 <- ifelse(grepl("^Z57", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z57)
    # Check for any value that starts with "Z59" and update Z59 column
    CHO_Roster2$Zcd_Z59 <- ifelse(grepl("^Z59", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z59)
    # Check for any value that starts with "Z60" and update Z60 column
    CHO_Roster2$Zcd_Z60 <- ifelse(grepl("^Z60", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z60)
    # Check for any value that starts with "Z62" and update Z62 column
    CHO_Roster2$Zcd_Z62 <- ifelse(grepl("^Z62", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z62)
    # Check for any value that starts with "Z63" and update Z63 column
    CHO_Roster2$Zcd_Z63 <- ifelse(grepl("^Z63", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z63)
    # Check for any value that starts with "Z64" and update Z64 column
    CHO_Roster2$Zcd_Z64 <- ifelse(grepl("^Z64", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z64)
    # Check for any value that starts with "Z65" and update Z65 column
    CHO_Roster2$Zcd_Z65 <- ifelse(grepl("^Z65", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z65)
    # Check for any value that starts with "Z75" and update Z75 column
    CHO_Roster2$Zcd_Z75 <- ifelse(grepl("^Z75", CHO_Roster2[, col]), 1, CHO_Roster2$Zcd_Z75)
  }
}

# Group by AHCCCSID and check for the presence of 1 in each Zcd column
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(across(starts_with("Zcd"), ~ as.integer(any(. == 1))))

# Create a new column with the sum across columns 2 to 11
collapsed_data$Sum_ZCode <- rowSums(collapsed_data[, 2:11], na.rm = TRUE)

# Create an indicator for if any Z codes are present across all diagnoses
collapsed_data <-
  collapsed_data |> 
  mutate(ZCodePresent = if_else(Sum_ZCode > 0, 1, 0))

# Select the first instance of AHCCCSID with priority to ZCode equals 1
selected_data <- collapsed_data |>
  arrange(AHCCCSID, desc(ZCodePresent)) |>
  distinct(AHCCCSID, .keep_all = TRUE)

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = selected_data,
        by = "AHCCCSID",
        all.x = TRUE)

Checkit <- CHO_Roster1 |> select(1, 36:59)

```

#### Homelessness

```{r}
#| label: Z Code HOmelessness
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Create a new column 'Homelessness' with default value 0
CHO_Roster2$Homelessness <- 0

# Check each column for the homelessness values 'Z59.00, Z59.01, and Z59.02'
for (col in c("PrimaryDiagnosis", "Dx1", "Dx2", "DX3", "Dx4", "Dx5", "DX6", "Dx7", "Dx8", "DX9", "Dx10", "Dx11", "DX12")) {
  CHO_Roster2$Homelessness[CHO_Roster2[[col]] == c("Z59.00", "Z59.01", "Z59.02")] <- 1
}

# Group by AHCCCSID and check for the presence of 1 in Homelessness column
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(Homelessness = as.integer(any(Homelessness == 1)))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

Checkit <- CHO_Roster1 |> select(1, 36:60)

```

### Inpatient Services

Indicators according to NCQA MY 2024 standards for Follow Up after Hospitalization for Mental Illness (FUH) values sets. These include the Inpatient Stay and Nonacute Inpatient value sets.

```{r}
#| label: Inpatient stay
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Inpatient Stay and Nonacute Inpatient Stay List

# Filter the MyHEDIS2023_MedicationDirectory for "Inpatient Stay"
InpatientStayValueSet <- MyHEDIS2023_AllValueSetsToCodes |>
  mutate_all(~str_trim(., side = "right")) |> # trim off extra spaces
  filter(`Value Set Name` == "Inpatient Stay")

# Extract the associated code values as a vector
InpatientStayValueSetList <- InpatientStayValueSet$Code

# Filter the MyHEDIS2023_MedicationDirectory for "Inpatient Stay"
NonacuteInpatientValueSet <- MyHEDIS2023_AllValueSetsToCodes |>
  mutate_all(~str_trim(., side = "right")) |> # trim off extra spaces
  filter(`Value Set Name` == "Nonacute Inpatient")

# Extract the associated code values as a vector
NonacuteInpatientValueSetList <- NonacuteInpatientValueSet$Code

# create the indicator column
CHO_Roster2 <-
  CHO_Roster2 |> 
  mutate(Inpatient = if_else(Svccode %in% InpatientStayValueSetList | Svccode %in% NonacuteInpatientValueSetList, 1, 0))

# Group by AHCCCSID and check for the presence of 1 in Inpatient column and collapse data so there is only one row per AHCCCSID
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(Inpatient = as.integer(any(Inpatient == 1)))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

Checkit <- CHO_Roster1 |> select(1, 36:61)

```

#### Cost: Inpatient Services

This assesses the cost of a given members inpatient stays for all inpatient and nonacute inpatient values.

```{r}
#| label: Cost Inpatient stay
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Calculate count, sum, and avg of IP services (all values)
CalcnetPd_IP <-
  CHO_Roster2 %>%
  mutate(Calcnetpd = as.numeric(Calcnetpd)) %>%
  filter(Svccode %in% c(InpatientStayValueSetList, NonacuteInpatientValueSetList)) %>%
  group_by(AHCCCSID) %>%
  summarize(count_IP = sum(!is.na(Calcnetpd)),
            total_calcnetpd_IP = sum(Calcnetpd),
            avg_calcnetpd_IP = mean(Calcnetpd)) |> 
  ungroup() %>%
  select(AHCCCSID,
         count_IP,
         total_calcnetpd_IP,
         avg_calcnetpd_IP)

# Add count, sum, and avg of IP services to CHO
CHO_Roster1 <-
  CHO_Roster1 %>%
  left_join(CalcnetPd_IP, by = "AHCCCSID") 

# Calculate count, sum, and avg of IP services if they are 0124 or 0126
CalcnetPd_0124 <-
  CHO_Roster2 %>%
  mutate(Calcnetpd = as.numeric(Calcnetpd)) %>%
  filter(Svccode %in% c("0124", "0126")) %>%
  group_by(AHCCCSID) %>%
  summarize(count_0124 = sum(!is.na(Calcnetpd)),
            total_calcnetpd_0124 = sum(Calcnetpd),
            avg_calcnetpd_0124 = mean(Calcnetpd)) |> 
  ungroup() %>%
  select(AHCCCSID,
         count_0124,
         total_calcnetpd_0124,
         avg_calcnetpd_0124)

# Add sum and avg of IP services to CHO
CHO_Roster1 <-
  CHO_Roster1 %>%
  left_join(CalcnetPd_0124, by = "AHCCCSID") 

Checkit <- CHO_Roster1 |> select(1, 36:67)

```

### Authorized Inpatient Service

This procedure utilizes the Alliance Daily Inpatient Census, which is a daily record obtained from BCBSAZ Health Choice indicating all individuals who are currently inpatient, or who were inpatient and have been discharged. This data is based on authorizations from HC medical management. The Inpatient Census includes cases with admission dates between 2023-01-01 and 2023-12-31.

#### All authorized inpatient stays

This will provide an indicator to the CHO Roster for any case that has an inpatient admission date that was included in the authorized inpatient service data provided by HCA.

```{r}
#| label: Inpatient Census
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Identify all cases on CHO Roster 1 that have an authorized inpatient admission
CHO_Roster1 <-
  CHO_Roster1 |> 
  mutate(Auth_IPAdmit = if_else(AHCCCSID %in% InpatientCensus$AHCCCSID, 1, 0))

# Count the number of admissions for each AHCCCS ID on the Inpatient Census

# Create a summary table grouping by AHCCCSID that counts the number of authorized inpatient admissions
summary_table <- 
  InpatientCensus |>
  group_by(AHCCCSID) |>
  summarise(Sum_Auth_IPAdmits = n())

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = summary_table,
        by = "AHCCCSID",
        all.x = TRUE)

Checkit <- CHO_Roster1 |> select(1, 39:69)

```

#### Maximum Length of Stay (LOS)

This procedure will calculate an individuals length of stay in an authorized inpatient stay. For those cases that do not have a discharge date, the discharge date will be recorded as the last day of the Measurement Year, 2023-12-31. For cases where there was more than 1 admission, the greatest length of stay will be recorded.

```{r}
#| label: Max Length of Stay
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

InpatientCensus <-
  InpatientCensus |> 
  mutate(DischargeDate = if_else(is.na(DischargeDate), "2023-12-31", DischargeDate)) |> #assigns 12/31/2023 to any cases who had not discharged as of the end of the 2023 calendar year to coincide with this range of data
  mutate(AdmissionDate = as.Date(AdmissionDate),
         DischargeDate = as.Date(DischargeDate))

# Calculate length of stay from the admit and discharge dates
summary_table <-
  InpatientCensus |> 
  mutate(LOS = DischargeDate - AdmissionDate) |> 
  select(AHCCCSID, LOS)

# Group by AHCCCSID, arrange by LOS in descending order, and select the first row (max) for each group
max_los_table <- 
  summary_table |>
  group_by(AHCCCSID) |>
  arrange(desc(LOS)) |>
  slice(1) |> 
  rename(Max_LOS = "LOS")

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = max_los_table,
        by = "AHCCCSID",
        all.x = TRUE)

# Group by AHCCCSID and calculate the average LOS
average_los_table <- 
  summary_table |>
  group_by(AHCCCSID) |>
  summarise(AVG_LOS = mean(LOS, na.rm = TRUE))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = average_los_table,
        by = "AHCCCSID",
        all.x = TRUE)

Checkit <- CHO_Roster1 |> select(1, 36:71)

```

#### Discharge Dispositions

*This is a topic that we can explore further. The data is somewhat irregular and would require interpretation. For now it is not being included*

### Emergency Department Visits 

Indicator according to NCQA MY 2024 standards for Follow Up after Emergency Department visit for Mental Illness (FUM) values sets.

#### Emergency Department (ED) Value Set

```{r}
#| label: FUM ED
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Inpatient Stay and Nonacute Inpatient Stay List

# Filter the MyHEDIS2023_MedicationDirectory for "ED"
EDValueSet <- MyHEDIS2023_AllValueSetsToCodes |>
  mutate_all(~str_trim(., side = "right")) |> # trim off extra spaces
  filter(`Value Set Name` == "ED")

# Extract the associated code values as a vector
EDValueSetList <- EDValueSet$Code

# create the indicator column
CHO_Roster2 <-
  CHO_Roster2 |> 
  mutate(ED = if_else(Svccode %in% EDValueSetList, 1, 0))

# Group by AHCCCSID and check for the presence of 1 in Inpatient column
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(ED = as.integer(any(ED == 1)))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

Checkit <- CHO_Roster1 |> select(1, 36:72)
```

##### Cost:Emergency Department

```{r}
#| label: Cost ED
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Calculate sum and avg of ED services
CalcnetPd_ED <-
  CHO_Roster2 %>%
  mutate(Calcnetpd = as.numeric(Calcnetpd)) %>%
  filter(Svccode %in% c(EDValueSetList)) %>%
  group_by(AHCCCSID) %>%
  summarize(count_ED = sum(!is.na(Calcnetpd)),
            total_calcnetpd_ED = sum(Calcnetpd), 
            avg_calcnetpd_ED = mean(Calcnetpd)) %>%
  ungroup() |> 
  select(AHCCCSID,
         count_ED,
         total_calcnetpd_ED,
         avg_calcnetpd_ED)

# Add sum and avg of IP services to CHO
CHO_Roster1 <-
  CHO_Roster1 %>%
  left_join(CalcnetPd_ED, by = "AHCCCSID") 

Checkit <- CHO_Roster1 |> select(1, 36:75)

```


#### Mental Illness and Intentional Self-harm (MISH) Value Set

```{r}
#| label: FUM MISH
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Filter the MyHEDIS2023_MedicationDirectory for "Mental Illness and Intentional Self-Harm "
MISHValueSet <- MyHEDIS2023_AllValueSetsToCodes |>
  mutate_all(~str_trim(., side = "right")) |> # trim off extra spaces
  filter(`Value Set Name` == "Mental Illness and Intentional Self-Harm")

# Extract the associated code values as a vector
MISHValueSetList <- MISHValueSet$Code

# Create a new column 'Homelessness' with default value 0
CHO_Roster2$MISH <- 0

# Check each column for the homelessness values 'Z59.00, Z59.01, and Z59.02'
for (col in c("PrimaryDiagnosis", "Dx1", "Dx2", "Dx3", "Dx4", "Dx5", "Dx6", "Dx7", "Dx8", "Dx9", "Dx10", "Dx11", "Dx12")) {
  CHO_Roster2$MISH[CHO_Roster2[[col]] %in% MISHValueSetList] <- 1
}

# Group by AHCCCSID and check for the presence of 1 in Inpatient column
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(MISH = as.integer(any(MISH == 1)))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

Checkit <- CHO_Roster1 |> select(1, 36:76)

```

```{r}
#| label: Cost MISH
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Calculate sum and avg of AOD services
CalcnetPd_MISH <- CHO_Roster2 %>%
  mutate(Calcnetpd = as.numeric(Calcnetpd)) %>%
  filter(
    PrimaryDiagnosis %in% MISHValueSetList |
    Dx1 %in% MISHValueSetList |
    Dx2 %in% MISHValueSetList |
    Dx3 %in% MISHValueSetList |
    Dx4 %in% MISHValueSetList |
    Dx5 %in% MISHValueSetList |
    Dx6 %in% MISHValueSetList |
    Dx7 %in% MISHValueSetList |
    Dx8 %in% MISHValueSetList |
    Dx9 %in% MISHValueSetList |
    Dx10 %in% MISHValueSetList |
    Dx11 %in% MISHValueSetList |
    Dx12 %in% MISHValueSetList
  ) %>%
  group_by(AHCCCSID) %>%
  summarize(
    count_MISH = sum(!is.na(Calcnetpd)),
    total_calcnetpd_MISH = sum(Calcnetpd), 
    avg_calcnetpd_MISH = mean(Calcnetpd)
  ) %>%
  ungroup() %>%
  select(
    AHCCCSID,
    count_MISH,
    total_calcnetpd_MISH,
    avg_calcnetpd_MISH
  )

# Add sum and avg of IP services to CHO
CHO_Roster1 <-
  CHO_Roster1 %>%
  left_join(CalcnetPd_MISH, by = "AHCCCSID") 

Checkit <- CHO_Roster1 |> select(1, 36:79)

```

### Antidepressant Medication 

Indicator according to NCQA MY 2024 standards for Antidepressant Medication Management (AMM) values sets.

```{r}
#| label: Antidepressant Medications
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# AMM List

# Filter the MyHEDIS2023_MedicationDirectory for "AMM"
AntidressantMedicationsValueSet <- MyHEDIS2023_MedicationDirectory |>
  mutate_all(~str_trim(., side = "right")) |> # trim off extra spaces
  filter(`Medication List Name` == "Antidepressant Medications")

# Extract the associated code values as a vector
AntidressantMedicationsValueSetList <- AntidressantMedicationsValueSet$Code

# create the indicator column
CHO_Roster2 <-
  CHO_Roster2 |> 
  mutate(AM = if_else(MedicationNDC %in% AntidressantMedicationsValueSetList, 1, 0))

# Group by AHCCCSID and check for the presence of 1 in Inpatient column
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(AM = as.integer(any(AM == 1)))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

Checkit <- CHO_Roster1 |> select(1, 36:80)

```

#### Cost: Antidepressant Medication

```{r}
#| label: Cost AM
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Calculate sum and avg of am services
CalcnetPd_AM <-
  CHO_Roster2 %>%
  mutate(Calcnetpd = as.numeric(Calcnetpd)) %>%
  filter(MedicationNDC %in% c(AntidressantMedicationsValueSetList)) %>%
  group_by(AHCCCSID) %>%
  summarize(count_AM = sum(!is.na(Calcnetpd)),
            total_calcnetpd_AM = sum(Calcnetpd), 
            avg_calcnetpd_AM = mean(Calcnetpd)) %>%
  ungroup() |> 
  select(AHCCCSID,
         count_AM,
         total_calcnetpd_AM,
         avg_calcnetpd_AM)

# Add sum and avg of IP services to CHO
CHO_Roster1 <-
  CHO_Roster1 %>%
  left_join(CalcnetPd_AM, by = "AHCCCSID") 

Checkit <- CHO_Roster1 |> select(1, 36:83)

```

### Substance Use and Dependance

Now add an indicator for Substance Abuse and Dependence according to NCQA MY 2024 standards for AOD Abuse and Dependence Value Set values sets.

#### Alchohol Abuse and Dependence

```{r}
#| label: AOD Substance Use diagnoses
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# AOD Abuse and Dependence  Stay List

# Filter the MyHEDIS2023_MedicationDirectory for "ED"
AODAbuseandDependenceValueSet <- MyHEDIS2023_AllValueSetsToCodes |>
  mutate_all(~str_trim(., side = "right")) |> # trim off extra spaces
  filter(`Value Set Name` == "AOD Abuse and Dependence")

# Extract the associated code values as a vector
AODAbuseandDependenceValueSetList <- AODAbuseandDependenceValueSet$Code

# Create a new column 'Homelessness' with default value 0
CHO_Roster2$AOD <- 0

# Check each column for the homelessness values 'Z59.00, Z59.01, and Z59.02'
for (col in c("PrimaryDiagnosis", "Dx1", "Dx2", "Dx3", "Dx4", "Dx5", "Dx6", "Dx7", "Dx8", "Dx9", "Dx10", "Dx11", "Dx12")) {
  CHO_Roster2$AOD[CHO_Roster2[[col]] %in% AODAbuseandDependenceValueSetList] <- 1
}

# Group by AHCCCSID and check for the presence of 1 in Inpatient column
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(AOD = as.integer(any(AOD == 1)))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

Checkit <- CHO_Roster1 |> select(1, 36:84)
```

##### Cost: Alcohol Abuse and Dependence

```{r}
#| label: Cost AOD
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Calculate sum and avg of AOD services
CalcnetPd_AOD <- CHO_Roster2 %>%
  mutate(Calcnetpd = as.numeric(Calcnetpd)) %>%
  filter(
    PrimaryDiagnosis %in% AODAbuseandDependenceValueSetList |
    Dx1 %in% AODAbuseandDependenceValueSetList |
    Dx2 %in% AODAbuseandDependenceValueSetList |
    Dx3 %in% AODAbuseandDependenceValueSetList |
    Dx4 %in% AODAbuseandDependenceValueSetList |
    Dx5 %in% AODAbuseandDependenceValueSetList |
    Dx6 %in% AODAbuseandDependenceValueSetList |
    Dx7 %in% AODAbuseandDependenceValueSetList |
    Dx8 %in% AODAbuseandDependenceValueSetList |
    Dx9 %in% AODAbuseandDependenceValueSetList |
    Dx10 %in% AODAbuseandDependenceValueSetList |
    Dx11 %in% AODAbuseandDependenceValueSetList |
    Dx12 %in% AODAbuseandDependenceValueSetList
  ) %>%
  group_by(AHCCCSID) %>%
  summarize(
    count_AOD = sum(!is.na(Calcnetpd)),
    total_calcnetpd_AOD = sum(Calcnetpd), 
    avg_calcnetpd_AOD = mean(Calcnetpd)
  ) %>%
  ungroup() %>%
  select(
    AHCCCSID,
    count_AOD,
    total_calcnetpd_AOD,
    avg_calcnetpd_AOD
  )

# Add sum and avg of IP services to CHO
CHO_Roster1 <-
  CHO_Roster1 %>%
  left_join(CalcnetPd_AOD, by = "AHCCCSID") 

Checkit <- CHO_Roster1 |> select(1, 50:87)

```

#### Unintentional Overdose

```{r}
#| label: OD Substance USe
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Unintentional Drug Overdose Stay List

# Filter the MyHEDIS2023_MedicationDirectory for "Unintentional Drug Overdose"
UnintentionalDrugOverdoseValueSet <- MyHEDIS2023_AllValueSetsToCodes |>
  mutate_all(~str_trim(., side = "right")) |> # trim off extra spaces
  filter(`Value Set Name` == "Unintentional Drug Overdose")

# Extract the associated code values as a vector
UnintentionalDrugOverdoseValueSetList <- UnintentionalDrugOverdoseValueSet$Code

# Create a new column 'Homelessness' with default value 0
CHO_Roster2$UOD <- 0

# Check each column for the homelessness values 'Z59.00, Z59.01, and Z59.02'
for (col in c("PrimaryDiagnosis", "Dx1", "Dx2", "Dx3", "Dx4", "Dx5", "Dx6", "Dx7", "Dx8", "Dx9", "Dx10", "Dx11", "Dx12")) {
  CHO_Roster2$UOD[CHO_Roster2[[col]] %in% UnintentionalDrugOverdoseValueSetList] <- 1
}

# Group by AHCCCSID and check for the presence of 1 in Inpatient column
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(UOD = as.integer(any(UOD == 1)))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

Checkit <- CHO_Roster1 |> select(1, 50:88)
```

##### Cost: Unintentional overdose

```{r}
#| label: Cost AOD
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Calculate sum and avg of AOD services
CalcnetPd_UOD <- CHO_Roster2 %>%
  mutate(Calcnetpd = as.numeric(Calcnetpd)) %>%
  filter(
    PrimaryDiagnosis %in% UnintentionalDrugOverdoseValueSetList |
    Dx1 %in% UnintentionalDrugOverdoseValueSetList |
    Dx2 %in% UnintentionalDrugOverdoseValueSetList |
    Dx3 %in% UnintentionalDrugOverdoseValueSetList |
    Dx4 %in% UnintentionalDrugOverdoseValueSetList |
    Dx5 %in% UnintentionalDrugOverdoseValueSetList |
    Dx6 %in% UnintentionalDrugOverdoseValueSetList |
    Dx7 %in% UnintentionalDrugOverdoseValueSetList |
    Dx8 %in% UnintentionalDrugOverdoseValueSetList |
    Dx9 %in% UnintentionalDrugOverdoseValueSetList |
    Dx10 %in% UnintentionalDrugOverdoseValueSetList |
    Dx11 %in% UnintentionalDrugOverdoseValueSetList |
    Dx12 %in% UnintentionalDrugOverdoseValueSetList
  ) %>%
  group_by(AHCCCSID) %>%
  summarize(
    count_UOD = sum(!is.na(Calcnetpd)),
    total_calcnetpd_UOD = sum(Calcnetpd), 
    avg_calcnetpd_UOD = mean(Calcnetpd)
  ) %>%
  ungroup() %>%
  select(
    AHCCCSID,
    count_UOD,
    total_calcnetpd_UOD,
    avg_calcnetpd_UOD
  )

# Add sum and avg of IP services to CHO
CHO_Roster1 <-
  CHO_Roster1 %>%
  left_join(CalcnetPd_UOD, by = "AHCCCSID") 

Checkit <- CHO_Roster1 |> select(1, 60:87)

```

### Major Depressive Disorder

```{r}
#| label: MDD
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Major Depressive Disorder

# Filter the MyHEDIS2023_MedicationDirectory for "MDD"
MajorDepressionValueSet <- MyHEDIS2023_AllValueSetsToCodes |>
  mutate_all(~str_trim(., side = "right")) |> # trim off extra spaces
  filter(`Value Set Name` == "Major Depression")

# Extract the associated code values as a vector
MajorDepressionValueSetList <- MajorDepressionValueSet$Code

# Create a new column 'Homelessness' with default value 0
CHO_Roster2$MDD <- 0

# Check each column for the homelessness values 'Z59.00, Z59.01, and Z59.02'
for (col in c("PrimaryDiagnosis", "Dx1", "Dx2", "Dx3", "Dx4", "Dx5", "Dx6", "Dx7", "Dx8", "Dx9", "Dx10", "Dx11", "Dx12")) {
  CHO_Roster2$MDD[CHO_Roster2[[col]] %in% MajorDepressionValueSetList] <- 1
}


# Group by AHCCCSID and check for the presence of 1 in Inpatient column
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(MDD = as.integer(any(MDD == 1)))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

Checkit <- CHO_Roster1 |> select(1, 70:92)
```

#### Cost: Major Depressive Disorder

```{r}
#| label: Cost AOD
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Calculate sum and avg of AOD services
CalcnetPd_MDD <- CHO_Roster2 %>%
  mutate(Calcnetpd = as.numeric(Calcnetpd)) %>%
  filter(
    PrimaryDiagnosis %in% MajorDepressionValueSetList |
    Dx1 %in% MajorDepressionValueSetList |
    Dx2 %in% MajorDepressionValueSetList |
    Dx3 %in% MajorDepressionValueSetList |
    Dx4 %in% MajorDepressionValueSetList |
    Dx5 %in% MajorDepressionValueSetList |
    Dx6 %in% MajorDepressionValueSetList |
    Dx7 %in% MajorDepressionValueSetList |
    Dx8 %in% MajorDepressionValueSetList |
    Dx9 %in% MajorDepressionValueSetList |
    Dx10 %in% MajorDepressionValueSetList |
    Dx11 %in% MajorDepressionValueSetList |
    Dx12 %in% MajorDepressionValueSetList
  ) %>%
  group_by(AHCCCSID) %>%
  summarize(
    count_MDD = sum(!is.na(Calcnetpd)),
    total_calcnetpd_MDD = sum(Calcnetpd), 
    avg_calcnetpd_MDD = mean(Calcnetpd)
  ) %>%
  ungroup() %>%
  select(
    AHCCCSID,
    count_MDD,
    total_calcnetpd_MDD,
    avg_calcnetpd_MDD
  )

# Add sum and avg of IP services to CHO
CHO_Roster1 <-
  CHO_Roster1 %>%
  left_join(CalcnetPd_MDD, by = "AHCCCSID") 

Checkit <- CHO_Roster1 |> select(1, 80:95)

```

### BIPOC

```{r}
#| label: BIPOC and Ethnicity
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Black, Indigenous, and People of Color
CHO_Roster1 <- CHO_Roster1 |> 
  mutate(Race_BIPOC = if_else(RaceCode != "C" & RaceCode != "7", 1, 0))

# Native American
CHO_Roster1 <- CHO_Roster1 |> 
  mutate(Race_NativeAmerican = if_else(RaceCode == "G", 1, 0))

# Hispanic (any)
CHO_Roster1 <- CHO_Roster1 |> 
  mutate(Ethnicity_HispanicAll = if_else(EthnicityCode == "1" | # Mexican, Mexican American, Chicano/a
                               EthnicityCode == "4" | # Another Hispanic, Latino or Spanish origin
                               EthnicityCode == "5"   # Hispanic or Latino Unknown  
                               , 1, 0))

# Mexican 
CHO_Roster1 <- CHO_Roster1 |> 
  mutate(Ethnicity_Mexican = if_else(EthnicityCode == "1", 1, 0))



Checkit <- CHO_Roster1 |> select(1, 80:99)
```

### Crisis Services

```{r}
#| label: Crisis Services
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

CrisisServices <- 
  data.frame(
    Code = c("H2011", "S9484", "S9485"),
    Description = c("Crisis intervention service, per 15 minutes", 
                    "Crisis intervention mental health service, per hour", 
                    "Crisis intervention mental health service, per diem")
  )

# All Crisis Services
# create the indicator column
CHO_Roster2 <-
  CHO_Roster2 |> 
  mutate(Crisis_all = if_else(Svccode %in% CrisisServices$Code, 1, 0))

# Group by AHCCCSID and check for the presence of 1 in Inpatient column
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(Crisis_all = as.integer(any(Crisis_all == 1)))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

# Community Based
# create the indicator column
CHO_Roster2 <-
  CHO_Roster2 |> 
  mutate(Crisis_Community = if_else(Svccode == "H2011", 1, 0))

# Group by AHCCCSID and check for the presence of 1 in Inpatient column
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(Crisis_Community = as.integer(any(Crisis_Community == 1)))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

# Facility Based
# create the indicator column
CHO_Roster2 <-
  CHO_Roster2 |> 
  mutate(Crisis_Facility = if_else(Svccode %in% c("S9484", "S9485"), 1, 0))

# Group by AHCCCSID and check for the presence of 1 in Inpatient column
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(Crisis_Facility = as.integer(any(Crisis_Facility == 1)))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

Checkit <- CHO_Roster1 |> select(1, 80:102)
```

### Peer and Family Run Organization (PFro) Services

```{r}
#| label: PFro Services
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

PFroProviders <- read.csv("./documents/PFroProviders.csv")

# create the indicator column
CHO_Roster2 <- 
  CHO_Roster2  |> 
  mutate(PFro = if_else(Fedtaxid %in% PFroProviders$ProviderTIN, 1, 0))

# Group by AHCCCSID and check for the presence of 1 in Inpatient column
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(PFro = as.integer(any(PFro == 1)))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)

Checkit <- CHO_Roster1 |> select(1, 80:103)
```

### Marriage, is what bwings us togevah today

```{r}
#| label: Marriage
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Create variables for Married and Divorced
CHO_Roster1 <-
  CHO_Roster1 |> 
  mutate(Married = if_else(MaritalStatus == "Married", 1, 0),
         DivorcedOrSeparated = if_else(MaritalStatus %in% c("Divorced", "Legally Separated"), 1, 0)
         )

Checkit <- CHO_Roster1 |> select(1, 80:105)
```

### Employment

```{r}
#| label: Employment
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Employment Status variables
CHO_Roster1 <-
  CHO_Roster1 |> 
  mutate(Employed = if_else(EmploymentStatus %in% c("Competitively Employed Part Time", "Competitively Employed Full Time"), 1,0),
         Employed_FT = if_else(EmploymentStatus == "Competitively Employed Full Time", 1, 0),
         Employed_PT = if_else(EmploymentStatus == "Competitively Employed Part Time", 1, 0)
  )


# Employment Services
# create the indicator column
CHO_Roster2 <-
  CHO_Roster2 |> 
  mutate(EmploymentServices = if_else(Svccode %in% c("H2025", "H2027"), 1, 0))

# Group by AHCCCSID and check for the presence of 1 in Inpatient column
collapsed_data <- CHO_Roster2 |>
  group_by(AHCCCSID) |>
  summarise(EmploymentServices = as.integer(any(EmploymentServices == 1)))

CHO_Roster1 <-
  merge(x = CHO_Roster1,
        y = collapsed_data,
        by = "AHCCCSID",
        all.x = TRUE)


Checkit <- CHO_Roster1 |> select(1, 80:109)

```

### Spoken Language

```{r}
#| label: Language
#| eval: true
#| include: false
#| echo: false
#| warning: false
#| error: false

# Language Variable
CHO_Roster1 <-
  CHO_Roster1 |> 
  mutate(EnglishSpeaking = if_else(SpokenLanguage == "ENGLISH", 1, 0),
         SpanishSpeaking = if_else(SpokenLanguage == "SPANISH", 1, 0),
         NonEnglishSpeaking = if_else(SpokenLanguage != "ENGLISH", 1, 0)
  )

Checkit <- CHO_Roster1 |> select(1, 80:112)
```

```{r}

write.csv(CHO_Roster1, "./data/output/CHO_Roster1.csv")

```









